---
title: "DRAFT figures for grad symposium 2020"
author: "Hanusia Higgins"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
stand_info <- read.csv("STAND_INFO_EAB_project_2020_cleaned_data.csv")
plot_info <- read.csv("PLOT_INFO_EAB_project_2020_cleaned_data.csv")
overstory_data <- read.csv("TREES_EAB_project_2020_cleaned_data.csv")
seedling_data <- read.csv("SEEDLINGS_EAB_project_2020_cleaned_data.csv")
small_sapling_data <- read.csv("SAPLINGS_SMALL_EAB_project_2020_cleaned_data.csv")
large_sapling_data <- read.csv("SAPLINGS_LARGE_EAB_project_2020_cleaned_data.csv")

library(tibble)
library(ggplot2)

#changing name of first column since it had some weird characters in it
names(overstory_data)[1] <- "plot_ID"
names(plot_info)[1] <- "plot_ID"
names(stand_info)[1]<- "Property"
names(seedling_data)[1] <- "plot_ID"
names(small_sapling_data)[1] <- "plot_ID"
names(large_sapling_data)[1] <- "plot_ID"
```

```{r BA dataframe setup, include=FALSE}
overstory_plus <- merge(x=overstory_data,y=plot_info,by="plot_ID")

#IN THE BELOW MODEL, I'm gonna convert things to use METRIC, not imperial, units!

overstory_plus$ASH_stump_DBH <- rep(NA)
overstory_plus$BA_sqm <- rep(NA)

#calculating BA for live trees & stumps using standard conversion from DBH to BA
#overstory_plus$BA_sqft <- overstory_plus$DBH_cm*overstory_plus$DBH_cm*0.005454

#BUT SINCE WE WANNA DO IT USING METRIC UNITS:
#RIGHT NOW THIS IS ONLY DOING IT CORRECTLY FOR LIVE TREES, NOT USING THE CORRECT DATA FOR STUMPS!!
overstory_plus$BA_sqm <- overstory_plus$DBH_cm*overstory_plus$DBH_cm*pi/40000

#only calculating the stump DBH w/ these constants for ash
for(i in nrow(overstory_plus)){
  if(overstory_plus$species[i]=="FRAM"){
    overstory_plus$ASH_stump_DBH[i] <- 
      (overstory_plus$DBH_cm[i]*((1.3716/overstory_plus$height_m[i])^-0.1074) +
         0.0685*(1.3716-overstory_plus$height_m[i]) + 0)
    # and then converting that to BA, for ASH stumps only
    overstory_plus$BA_sqm[i] <- 
      overstory_plus$ASH_stump_DBH[i]*overstory_plus$ASH_stump_DBH[i]*pi/40000 
  }
  
  plot_names <- unique(overstory_plus$plot_ID)
  ash_cut <- data.frame("plot_ID" = plot_info$plot_ID, 
                        "stand_ID" = plot_info$Stand_name,
                        "harvest_status" = plot_info$harvest_status, 
                        "gap_status" = plot_info$gap_status, 
                        "live_ash_sum"=rep(0), #starting @ zero so they can be added to
                        "cut_ash_sum"=rep(0) #ditto
                        #,"any_live_ash"=rep("NA"), #but adding a separate var to tell
                        #"any_cut_ash"=rep("NA") #whether it actually went thru /was counted
  )
  
  for (i in 1:nrow(overstory_plus)){ #iterating thru each tree/stump/snag in all plots
    
    #basically, adding the BA of each live ash tree in a plot to the sum value
    if(overstory_plus$species[i]=="FRAM"){
      if(overstory_plus$status[i]=="live"){
        ash_cut$live_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] <- 
          ash_cut$live_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] +
          overstory_plus$BA_sqm[i]
        #using BA calculated from DBH calculated w/ the formula I used above!
      } else if(overstory_plus$status[i]=="stump") {
        ash_cut$cut_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] <- 
          ash_cut$cut_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] +
          overstory_plus$BA_sqm[i]
      }
    }
    
  }
}
```


### First up, looking at ash basal area in meters (live and cut), per sample plot. 
Refer to code comments for details on data included in each individual figure.

```{r ash BA boxplots, include = TRUE}
#this one maps BA (in sq m) of standing ash against harvest status 
#across all plots
boxplot(formula=live_ash_sum~harvest_status, data=ash_cut)

#ditto this one, except comparing only "matrix" plots against unharvested
boxplot(formula=live_ash_sum~harvest_status, 
        data=ash_cut[ash_cut$gap_status=="NO",])

#now just looking @ the spread of CUT ash in harvested sites
boxplot(formula=cut_ash_sum~gap_status,
        data=ash_cut[ash_cut$harvest_status=="YES",])

#are they significant?
live_ash_BA_all <- aov(formula=live_ash_sum~harvest_status, data=ash_cut)
print(live_ash_BA_all)
print(summary(live_ash_BA_all)) #yes!

live_ash_BA_mat <- aov(formula=live_ash_sum~harvest_status, 
                       data=ash_cut[ash_cut$gap_status=="NO",])
print(live_ash_BA_mat)
print(summary(live_ash_BA_mat))
```

### Next, looking at ash seedling counts per stand.
Refer to code comments for details on data included in each individual plot.

```{r seedling dataframe setup, include=FALSE}
#merging the seedlings df w/ the plot_info df on the basis of the plot_ID column
seedlings_plus <- merge(x=seedling_data,y=plot_info,by="plot_ID")

ash_seedlings <- data.frame("plot_ID" = plot_info$plot_ID, 
                            "stand_ID" = plot_info$Stand_name,
                            "harvest_status" = plot_info$harvest_status, 
                            "gap_status" = plot_info$gap_status, 
                            "ash_seedling_sum"=rep(0) #starting @ zero so they can be added to
)

#for loop to sum up all the ash seedlings...
for(i in 1:nrow(seedlings_plus)){ #iterating thru each seedling!
  if(seedlings_plus$species[i]=="FRAM"){ #only looking @ ash seedlings for now
    #ash_cut$live_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] <- 
    #  ash_cut$live_ash_sum[ash_cut$plot_ID==overstory_plus$plot_ID[i]] +
    #  overstory_plus$BA_sqm[i]
    ash_seedlings$ash_seedling_sum[seedlings_plus$plot_ID[i]==ash_seedlings$plot_ID] <-
      ash_seedlings$ash_seedling_sum[seedlings_plus$plot_ID[i]==ash_seedlings$plot_ID] +
      seedlings_plus$tally[i]
  }
}

#now, breaking it down by STAND

ash_seedlings_stand <- data.frame("stand_ID" = stand_info$Stand_name, 
 #                           "harvest_status" = stand_info$Harvest_status, 
 # making harvest status a binary variable of cut vs. not (not implicating "to be cut")
                            "harvest_status" = ifelse(stand_info$Harvest_status=="cut",TRUE,FALSE) ,
                           "ash_seedling_sum"=rep(0) #starting @ zero so they can be added to
)

#for loop to sum up all the ash seedlings...
for(i in 1:nrow(ash_seedlings)){ #iterating thru each PLOT!
    ash_seedlings_stand$ash_seedling_sum[ash_seedlings$stand_ID[i]==ash_seedlings_stand$stand_ID] <-
      ash_seedlings_stand$ash_seedling_sum[ash_seedlings$stand_ID[i]==ash_seedlings_stand$stand_ID] +
      #seedlings_plus$tally[i]
      ash_seedlings$ash_seedling_sum[i]
}

```

```{r ash seedling figures}
# more boxplots! ash seedlings @ stand-level -------------------------------
#now to create some basic boxplots comparing ash seedlings
# in harvested and non-harvested sites
p10=boxplot(formula=ash_seedling_sum~harvest_status, data=ash_seedlings_stand)
print(p10)
#boxplot is NOT very useful as both are clustered around 0...
#maybe a "dot plot" would be better??
anova1 <- aov(formula=ash_seedling_sum~harvest_status, data=ash_seedlings_stand)
print(summary(anova1)) #this one IS significant.

#dev.off()
#let's try a dot plot (then maybe a density plot?)
p11 <-ggplot(data=ash_seedlings_stand, aes(x=harvest_status, y=ash_seedling_sum)) +
  geom_dotplot(binaxis='y', stackdir='center')
print(p11)

#now let's try a density plot:
p12 <-ggplot(data=ash_seedlings_stand, aes(x=ash_seedling_sum)) + # number of seedlings on x-axis
  geom_density(aes(fill=factor(harvest_status)), #separating by treatment
               alpha=0.6)
print(p12) #this still doesn't look great IMO...but probably better than the boxplot

```

### Finally, plotting ash seedling # against proportion of ash BA removed (as a proxy for harvest intensity, in harvested stands only):

Even when I remove the outliers, the regression line doesn't really show a trend here...

```{r seedlings vs. harvest intensity, echo=FALSE}
#do the same thing for cut/live ash by STAND
ash_cut_stand <- data.frame("stand_ID" = stand_info$Stand_name, 
        # making harvest status a binary variable of cut vs. not (not implicating "to be cut")
                      "harvest_status" = ifelse(stand_info$Harvest_status=="cut",TRUE,FALSE) ,
                      "live_ash_sum"=rep(0), #starting @ zero so they can be added to
                      "cut_ash_sum"=rep(0) #ditto
                      #,"any_live_ash"=rep("NA"), #but adding a separate var to tell
                      #"any_cut_ash"=rep("NA") #whether it actually went thru /was counted
)

#for loop to sum up harvested BA by stand...
for(i in 1:nrow(ash_cut)){ #iterating thru each PLOT!
  ash_cut_stand$live_ash_sum[ash_cut$stand_ID[i]==ash_cut_stand$stand_ID] <-
    ash_cut_stand$live_ash_sum[ash_cut$stand_ID[i]==ash_cut_stand$stand_ID] +
    ash_cut$live_ash_sum[i] #adding total BA per plot......
  ash_cut_stand$cut_ash_sum[ash_cut$stand_ID[i]==ash_cut_stand$stand_ID] <-
    ash_cut_stand$cut_ash_sum[ash_cut$stand_ID[i]==ash_cut_stand$stand_ID] +
    ash_cut$cut_ash_sum[i]
}

#adding proportion of ash BA cut:
ash_cut_stand$prop_BA_cut <- ash_cut_stand$cut_ash_sum /
  (ash_cut_stand$live_ash_sum+ash_cut_stand$cut_ash_sum)

#glimpse(ash_cut_stand)

#now to merge THESE two dataframes:
ash_seed_cut_stand <- merge(x=ash_seedlings_stand, y=ash_cut_stand, by="stand_ID")
glimpse(ash_seed_cut_stand)

#NOW to look at a regression with STAND-LEVEL data:
#looking @ harvested sites only:

p5 <- ggplot(data=ash_seed_cut_stand[ash_seed_cut_stand$harvest_status.x==TRUE,],
             aes(x=prop_BA_cut, y=ash_seedling_sum)) + 
  geom_point(color="#000099") #+
#  geom_smooth(method=lm) +
#  ylim(0,15) + #to crop that outlier 
#  xlim(0.2,1) #to crop a DIFFERENT outlier
p5 #there's a biiiig outlier.....


#with outliers dropped & regression line added:
p4 <- ggplot(data=ash_seed_cut_stand[ash_seed_cut_stand$harvest_status.x==TRUE,],
             aes(x=prop_BA_cut, y=ash_seedling_sum)) + 
  geom_point(color="#000099") +
  geom_smooth(method=lm) +
  ylim(0,15) + #to crop that outlier 
  xlim(0.2,1) #to crop a DIFFERENT outlier
p4 #there's a biiiig outlier.....
#max(ash_seed_cut_stand$ash_seedling_sum)
#summary(p4)


```



